### 2024-08-25 thesis/tex/abstract-eng.tex
docs: updated thesis template

\cleardoublepage

\secondabstract

Abstract

\secondkeywords

Keyword1, Keyword2, Keyword3