### 2024-08-25 thesis/tex/abstract-pl.tex

\cleardoublepage

\abstract

Test

\keywords

Keyword1, Keyword2, Keyword3