### 2024-10-02 thesis/tex/chapters/5-summary.tex
thesis: add initial structure of the chapters

\clearpage
\section{Podsumowanie}

**TODO** [ok. 10 stron]

\begin{enumerate}
    \item Najważniejsze cechy wytworzonego rozwiązania
    \item Plan rozbudowy
\end{enumerate}

    Bibliografia \cite{knuth1984texbook}.