### 2024-12-20 thesis/tex/abstract-eng.tex
thesis: reformat used tools subsection

\cleardoublepage{}

\secondabstract{}

Abstract

\secondkeywords{}

Keyword1, Keyword2, Keyword3