### 2024-12-20 thesis/tex/abstract-pl.tex

\cleardoublepage{}

\abstract{}

Test

\keywords{}

Keyword1, Keyword2, Keyword3