### 2024-12-27 thesis/tex/abstract-pl.tex
thesis: add missing pictures

\cleardoublepage{}

\abstract{}

Test

\keywords{}

Słowo kluczowe 1, słowo kluczowe 2, słowo kluczowe 3.