### 2025-01-15 thesis/tex/abstract-eng.tex
thesis: add keywords, small fixes to bibliography

\cleardoublepage{}

\secondabstract{}

Abstract

\secondkeywords{}
DNA sequences, taxonomic classification,  metagenomics, artificial neural networks, contrastive learning