### 2025-01-15 thesis/tex/abstract-pl.tex

\cleardoublepage{}

\abstract{}

Streszczenie

\keywords{}
sekwencje DNA, klasyfikacja taksonomiczna, metagenomika, sztuczne sieci neuronowe, uczenie kontrastowe